---
title: "Spotify Data"
author: "Kodi Obika and Eryk Pecyna"
date: "May 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
data <- read.csv("TrackData.csv")
```

## Introduction


This is an analysis of the data collected and calculated by Spotify regarding the Billboard Hot 100 songs for the years 1960-2015. The data was calculated using a variety of features and can be found described in more detail here <https://www.rdocumentation.org/packages/billboard/versions/0.1.0/topics/spotify_track_data>


## Plots of Spotify Data

```{r RawData, echo = FALSE}

ggplot(data, aes(x = danceability)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Danceability") +
  ggtitle("Danceability Frequency")

ggplot(data, aes(x = energy)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Energy") +
  ggtitle("Energy Frequency")

ggplot(data, aes(x = loudness)) + 
  geom_histogram(binwidth = 1, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Loudness") +
  ggtitle("Loudness Frequency")

ggplot(data, aes(x = speechiness)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Speechiness") +
  ggtitle("Speechiness Frequency")

ggplot(data, aes(x = acousticness)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Acousticness") +
  ggtitle("Acousticness Frequency")

ggplot(data, aes(x = instrumentalness)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Instrumentalness") +
  ggtitle("Instrumentalness Frequency")

ggplot(data, aes(x = liveness)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Liveness") +
  ggtitle("Liveness Frequency")

ggplot(data, aes(x = valence)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Valence") +
  ggtitle("Valence Frequency")

ggplot(data, aes(x = tempo)) + 
  geom_histogram(binwidth = 5, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Tempo") +
  ggtitle("Tempo Frequency")

ggplot(data, aes(x = duration_ms)) + 
  geom_histogram(binwidth = 20000, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Duration") +
  ggtitle("Duration Frequency")





```




Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
