---
title: "Spotify Data"
author: "Kodi Obika and Eryk Pecyna"
date: "May 7, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
data <- read.csv("TrackData.csv")
perm.test <- function(x, y, z, n) {
  mu.z = mean(x[y == z])
  mu.nz = mean(x[y != z])
  permutations = numeric(n)
  for (i in 1:n) {
    resample = sample(x)
    zs = resample[1:(length(x)/2)]
    nzs = resample[(length(x)/2+1):length(x)]
    permutations[i] = mean(zs) - mean(nzs)
  }
  hist(permutations, freq = FALSE)
  test.statistic = mu.z - mu.nz
  abline(v = test.statistic, lwd = 3)
  print("p-value:")
  print(2*mean(permutations < test.statistic))
  var.p = var(x) * 2 / (length(x)/2)
  curve(dnorm(x, mean = 0, sd = sqrt(var.p)), add = TRUE, lwd = 3)
}

clt <- function(y, n = 10000) {
  avg <- numeric(n)
  for (i in 1:n) {
    avg[i] <- mean(sample(y, 10))
  }
  hist(avg, probability = TRUE)
  curve(dnorm(x, mean(y), sd(y)/sqrt(10)), add = TRUE, col = "darkmagenta")
}


linreg <- function(xCol, yCol, xLabel = "X", yLable = "Y") {
  m1 <- rep(1, length(xCol))
  m2 <- xCol
  A <- cbind(m1, m2)
  B <- t(A)%*%A
  BInv <- solve(B)
  P <- A%*%BInv%*%t(A)
  Projection <- P%*%yCol
  plot(xCol, yCol, pch = ".", cex = 2)
  points(xCol, Projection, col = "darkmagenta", pch =".", cex = 3)
  R2 <- var(Projection)/var(yCol)
  print("Correlation:")
  print(R2[1])
  R2[1]
}

meancorrelation <- function(dt, testcol) {
  total <- 0
  for(i in 1:ncol(dt)) {
    if(i != testcol) {
      total <- total + linreg(dt[,testcol], dt[,i])
    }
  }
  print("Mean Correlation:")
  print(total / (ncol(dt) - 1))
  total / (ncol(dt) - 1)
}

PCA <- function(RawData, nV = 1) {
  numRow <- nrow(RawData)
  numCol <- length(RawData[1,])
  fillZeroes <- function(numCols) {
    matrix(0, nrow = numRow, ncol = numCols)
  }
  Adjusted <- scale(RawData, center=TRUE, scale = c(rep(sqrt(numRow-1), numCol)))
  S <- var(RawData)
  Eig <- eigen(S)
  P <- Eig$vectors
  PInv <- solve(P)
  Data.eig <- Adjusted%*%P
  A.reconstruct <- Data.eig%*%PInv
  Reconstructed <- scale(A.reconstruct, scale = c(rep(1/sqrt(numRow-1), numCol)))
  Reconstructed <- scale(Reconstructed, center = -colMeans(RawData), scale = FALSE)
  AStripped <- cbind(Data.eig[,1:nV], fillZeroes(numCol-nV))
  A.reconstruct <- AStripped%*%PInv
  Reconstructed <- scale(A.reconstruct, scale = c(rep(1/sqrt(numRow-1), numCol)))
  Reconstructed <- scale(Reconstructed, center = -colMeans(RawData), scale = FALSE)
  Reconstructed
}




```

## Introduction


This is an analysis of the data collected and calculated by Spotify regarding the Billboard Hot 100 songs for the years 1960-2015. The data was calculated using a variety of features and can be found described in more detail here <https://www.rdocumentation.org/packages/billboard/versions/0.1.0/topics/spotify_track_data>


## Plots of Spotify Data

```{r RawData}

ggplot(data, aes(x = danceability)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Danceability") +
  ggtitle("Danceability Frequency")

ggplot(data, aes(x = energy)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Energy") +
  ggtitle("Energy Frequency")

ggplot(data, aes(x = loudness)) + 
  geom_histogram(binwidth = 1, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Loudness") +
  ggtitle("Loudness Frequency")

ggplot(data, aes(x = speechiness)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Speechiness") +
  ggtitle("Speechiness Frequency")

ggplot(data, aes(x = acousticness)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Acousticness") +
  ggtitle("Acousticness Frequency")

ggplot(data, aes(x = instrumentalness)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Instrumentalness") +
  ggtitle("Instrumentalness Frequency")

ggplot(data, aes(x = liveness)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Liveness") +
  ggtitle("Liveness Frequency")

ggplot(data, aes(x = valence)) + 
  geom_histogram(binwidth = 0.025, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Valence") +
  ggtitle("Valence Frequency")

ggplot(data, aes(x = tempo)) + 
  geom_histogram(binwidth = 5, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Tempo") +
  ggtitle("Tempo Frequency")

ggplot(data, aes(x = duration_ms)) + 
  geom_histogram(binwidth = 20000, fill = "darkmagenta") +
  theme_minimal() +
  xlab("Duration") +
  ggtitle("Duration Frequency")

```

##Representation of each key
This displays the counts of songs that were written in each key.
```{r Chords}
keymode_lab <- c("C", "Cm", "C#/Db", "C#m/Dbm", "D", "Dm", "D#/Eb", 
                 "D#m/Ebm", "E", "Em", "F", "Fm", "F#/Gb", "F#m/Gbm", 
                 "G", "Gm", "G#/Ab", "G#m/Abm", "A", "Am", "A#/Bb", 
                 "A#m/Bbm", "B", "Bm")

data %>% 
  select(key, mode) %>% 
  group_by(key, mode) %>% 
  mutate(n = n()) %>% 
  unique() %>% 
  arrange(key) %>%
  ungroup() %>% 
  mutate(keymode = keymode_lab) %>% 
  select(keymode, n) %>% 
  ggplot(aes(x = keymode, y = n)) + 
  geom_bar(stat = "identity", fill = "darkmagenta") +
  coord_flip()
```


##Are danceable songs more explicit?

The P-value is tiny because it is extremely unlikely to observe such a difference by random chance. This means that explicit songs are much more likely to be danceable than non-explicit songs.  

```{r perms}
perm.test(data$danceability, data$explicit, FALSE, 10000)
```



##KODI RENAME THIS

```{r clt}
clt(data$danceability)
```


##Linear Regression 
Here we use a Projection Matrix approach to calculate linear regression
```{r linreg, fig.show="hide", results="hide"}
Attributes <- cbind(data$danceability, data$energy, data$loudness,
                    data$speechiness, data$acousticness, data$instrumentalness,
                    data$liveness, data$valence, data$tempo)

numericalData <- (cbind(data$Position, Attributes))  

N <- ncol(numericalData)
means <- numeric(N)
for(i in 1:N) {
  means[i] <- meancorrelation(numericalData, i)
}
```

The third column, which happens to be energy in our definition of numericalData, has the highest average correlation with the other columns. The following linear regressions are the highest correlations energy has with any other columns.
```{r means} 
means

linreg(data$energy, data$loudness)
linreg(data$energy, data$acousticness)
linreg(data$energy, data$valence)
```

##Principal Components Analysis
Here we will use Principal Componenets Analysis to see if we can reconstruct any of the columns after stripping some of the eigenvectors of the basis of eigenvectors formed from the covariance matrix.

Using one eigenvector gives us a good reconstruction of tempo (column 9)
Using two also gives a good reconstruction of loudness (column 2) but not much else
Using eight gives a decent reconstruction of all columns except speechiness (column 4)
Using nine yields a perfect reconstruction of all the data
```{r PCA}
head(PCA(Attributes)); head(Attributes)
head(PCA(Attributes, 2)); head(Attributes)
head(PCA(Attributes, 8)); head(Attributes)
head(PCA(Attributes, 9)); head(Attributes)
```




